Title

### 2012-01-13 inc/laser/laser.tex
import OE paper up to (not inc) subsection 3

To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique [16] to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal. 
For the employed non-Brewster crystal cut (Fig. 1(a)), a ray
%page 8633
transfer matrix analysis (with the $x$ direction parallel to the crystallographic $b$-axis and the $y$
direction parallel to the $a$-axis) reveals that the required ellipticity for a stigmatic thermal lens
generated by a parabolic temperature distribution, $T(x, y) = T_0 - \frac{1}{2} ( A_\smallT x^2 + B_\smallT y^2 )$, where $T_0$
is the on-axis temperature, is given by
%TODO make W symbol consistent with the rest of the paper
\begin{equation} \label{eq:laser_ratio_ba}
  \frac{B_\smallT}{A_\smallT} = 
  \frac{ \cos^2 \theta_2 \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) }
       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) \right] }
  \quad\text{.}
\end{equation}
Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 [11,15]), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
expansion with coefficient $\alpha T$ [16].
Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region [30].
In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
In other words, for the potassium gadolinium tungstate (\ce{KGd(WO4)2}) crystal geometry of Fig. 1(a), we expect from \ref{eq:laser_ratio_ba} to require $\dfrac{B_\smallT}{A_\smallT} \approx \dfrac{ \cos^2 \theta_2 }{ \cos^2 \theta_1 } = 1.3$.

The spatial pump distribution needed to generate the required $B_T/A_T$ ratio can be found using a cosine Fourier series solution to the thermal diffusion equation for our two-dimensional problem,
\begin{equation} \label{eq:heat_eqn}
  \kappa \nabla^2 T(x,y) = Q(x,y)
  \quad\text{,}
\end{equation}
where $\kappa$ is the thermal diffusion coefficient.
Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry [16].
The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
\begin{equation}
  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right]
  \quad\text{;}
\end{equation}
that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique [16]; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
This analytical result for the central region around the laser beam axis is shown in Fig. 2; displayed are the
%page 8634

>>>>

To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique [16] to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal. 
For the employed non-Brewster crystal cut (Fig. 1(a)), a ray
%page 8633
transfer matrix analysis (with the $x$ direction parallel to the crystallographic $b$-axis and the $y$
direction parallel to the $a$-axis) reveals that the required ellipticity for a stigmatic thermal lens
generated by a parabolic temperature distribution, $T(x, y) = T_0 - \frac{1}{2} ( A_\smallT x^2 + B_\smallT y^2 )$, where $T_0$
is the on-axis temperature, is given by
%TODO check W symbol consistent with the rest of the paper
\begin{equation} \label{eq:laser_ratio_ba}
  \frac{B_\smallT}{A_\smallT} = 
  \frac{ \cos^2 \theta_2 \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) }
       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) \right] }
  \quad\text{.}
\end{equation}
Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 [11,15]), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
expansion with coefficient $\alpha T$ [16].
Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region [30].
In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
In other words, for the potassium gadolinium tungstate (\ce{KGd(WO4)2}) crystal geometry of Fig. 1(a), we expect from \ref{eq:laser_ratio_ba} to require $\dfrac{B_\smallT}{A_\smallT} \approx \dfrac{ \cos^2 \theta_2 }{ \cos^2 \theta_1 } = 1.3$.

The spatial pump distribution needed to generate the required $B_T/A_T$ ratio can be found using a cosine Fourier series solution to the thermal diffusion equation for our two-dimensional problem,
\begin{equation} \label{eq:heat_eqn}
  \kappa \nabla^2 T(x,y) = Q(x,y)
  \quad\text{,}
\end{equation}
where $\kappa$ is the thermal diffusion coefficient.
Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry [16].
The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
\begin{equation}
  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right]
  \quad\text{;}
\end{equation}
that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique [16]; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
This analytical result for the central region around the laser beam axis is shown in Fig. 2; displayed are the
%page 8634
horizontal ($y = 0$, Fig. 2(a)) and vertical ($x = 0$, Fig. 2(b)) sections, respectively, through the $Q(x,y)$ and $T(x,y)$ distributions.
Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (Fig. 1(a)).
This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium [1]; in particular, in the vertical direction (Fig. 2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.

% insert figure 2

The analysis clearly shows that the TEM$_{00}$ laser mode will experience a near perfect pump-induced thermal lens upon propagation through the gain medium --- the deviation of $T(x,y)$ from the parabolic approximation, and hence a perfect dominant thermo-optic (GRIN) duct, being insignificant over the size of the laser mode in both the horizontal and vertical directions.
Moreover, the fact that $W_x \approx 1.6 w_x$ implies that some degree of horizontal spatial walk-off between the pump and laser beams can be tolerated in this TLS geometry.
For example, when the pump and laser beams are exactly overlapped at the center of the Yb:KGW crystal, an external angle of 0.1rad (= 5.7$^\circ$) would generate an effective spatial walk-off of only about 75$\mu$m over our 3mm gain crystal length.
Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium [1].

We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) [11,14] and stress-induced refractive index changes [31].
For the employed crystal cut in our TLS geometry (Fig. 1(a)), the anisotropy in $\kappa$ [14,27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
The refractive index
%page 8635
changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index [21].
Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
Consequently, for our laser polarized along the optical $N_p$-axis [5], there is insufficient information to include accurately the effects of stress-induced refractive index changes.

\subsection{The Yb:KGW laser cavity}
